Locally Conservative Discontinuous Petrov-Galerkin Finite Elements for Fluid Problems

### 2013-12-10 docs/Truman/LocalConservation/paper/paper.tex

\documentclass[letterpaper]{article}
%TODO: check for periods after equations
%TODO: bold face proper variables
%TODO: Make sure titles have consistent case
%TODO: Make sure figures work in gray-scale
\usepackage[margin=1.0in]{geometry}
\usepackage{authblk}
\usepackage{amsmath, amssymb, amsthm, mathtools}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{color}
\usepackage{subcaption}
\usepackage{morefloats}
\usepackage{pgf,tikz}
\usetikzlibrary{arrows}

@@

\begin{abstract}
We develop a locally conservative formulation of the discontinuous
Petrov-Galerkin finite element method (DPG) for convection-diffusion type
problems using Lagrange multipliers to exactly enforce conservation over each
element. We provide a proof of convergence as well as extensive numerical
experiments showing that the method is indeed locally conservative. We also
show that standard DPG, while not guaranteed to be conservative, is nearly
conservative for many of the benchmarks considered. The new method preserves many of
the attractive features of DPG, but turns the normally symmetric
positive-definite DPG system into a saddle point problem.
\end{abstract}

\section{Introduction}
The discontinuous Petrov-Galerkin (DPG) method with optimal test functions has
been under active development for convection-diffusion type systems
\cite{DPG1, DPG2, DPG3, DPG5, DemkowiczHeuer, ChanHeuerThanhDemkowicz2012,
MoroNguyenPeraire11}. In this paper, we develop a theory for the locally conservative
formulation of DPG for convection-diffusion type equations and supplement this
with extensive numerical results.

@@

The discontinuous Petrov-Galerkin finite element method has been described as
least squares finite elements with a twist. The key difference is that least
squares methods seek to minimize the residual of the solution in the $L^2$
norm, while DPG seeks the minimization in a dual norm realized through the
inverse Riesz map. Exact mass conservation has been an issue that has plagued
least squares finite elements for a long time. Several approaches have been
used to try to address this. Bochev \etal\cite{Bochev2010} accomplish local
conservation by using a pointwise divergence free velocity space in the Stokes
formulation.  Chang and Nelson\cite{ChangNelson1997} developed the
\emph{restricted LSFEM}\cite{ChangNelson1997} by augmenting the least squares
equations with a Lagrange multiplier explicitly enforcing mass conservation
element-wise. Our conservative formulation of DPG takes a similar approach and
both methods share similar negative of transforming a minimization method to a
saddle point problem. In the interest of crediting Chang and Nelson's
restricted LSFEM, we call the following locally conservative DPG method the
restricted DPG method (RDPG).

\subsection{DPG is a Minimum Residual Method}
Roberts \etal presents a brief history and derivation of DPG with optimal test functions in
\cite{DPGStokes}. We follow his derivation of the standard DPG method as a
minimum residual method. Let $U$ be the trial Hilbert space and $V$ the test
Hilbert space for a well-posed variational problem $b(u,v)=l(v)$. In operator
form this is $Bu=l$, where $B:U\rightarrow V'$ and $\LRa{Bu,v}=b(u,v)$. We seek to minimize the
residual for the discrete space $U_h\subset U$:
\begin{equation}
u_h=\argmin_{u_h\in U_h}\frac{1}{2}\norm{Bu_h-l}^2_{V'}\,.
\label{minresidual}
\end{equation}
Recalling that the Riesz operator $R_V:V\rightarrow V'$ is an isometry defined
by
\[
\LRa{R_Vv,\delta v}=\LRp{v,\delta v}_V,\quad\forall\delta v\in V,
\]
we can use the Riesz inverse to minimize in the $V$-norm rather than its dual:
\begin{equation}
\frac{1}{2}\norm{Bu_h-l}^2_{V'}=\frac{1}{2}\norm{R_V^{-1}(Bu_h-l)}^2_V
=\frac{1}{2}\LRp{R_V^{-1}(Bu_h-l),R_V^{-1}(Bu_h-l)}_V\,.
\label{eq:rieszapplied}
\end{equation}
The first order optimality condition for \eqnref{rieszapplied} requires
the G\^ateaux derivative to be zero in all directions $\delta u \in
U_h$, i.e.,
\[
\left(R_V^{-1}(Bu_h-l),R_V^{-1}B\delta u\right)_V = 0, \quad \forall \delta u \in U.
\]
By definition of the Riesz operator, this is equivalent to
\begin{equation}
\LRa{Bu_h-l,R_V^{-1}B\delta u_h}=0\quad\forall\delta u_h\in U_h\,.
\label{eq:DPGbilinearform}
\end{equation}
Now, we can identify $v_{\delta u_h}\coloneqq R_V^{-1}B\delta u_h$ as the
optimal test function for trial function $\delta u_h$. Define $T:=R_V^{-1}B:U_h\rightarrow V$ as the trial-to-test operator. Now we can rewrite
\eqnref{DPGbilinearform} as
\begin{equation}
b(u_h,v_{\delta u_h})=l(v_{\delta u_h}).
\label{eq:DPGmethod}
\end{equation}
The DPG method then is to solve \eqnref{DPGmethod} with optimal test functions
$v_{\delta u_h}\in V$ that solve the auxiliary problem
\begin{equation}
\LRp{v_{\delta u_h},\delta v}_V=\LRa{R_Vv_{\delta u_h},\delta v}
=\LRa{B\delta u_h,\delta v}=b(\delta u_h,\delta v)\,,\quad\forall\delta v\in V.
\label{eq:optimaltestproblem}
\end{equation}
Using a continuous test basis would result in a global solve for every optimal
test function. Therefore DPG uses a discontinuous test basis which makes each
solve element-local and much more computationally tractable. Of course,
\eqnref{optimaltestproblem} still requires the inversion of the
infinite-dimensional Riesz map, but approximating $V$ by a finite
dimensional $V_h$ which is of a higher polynomial degree than $U_h$ (hence
``enriched space'') works well in practice.

@@

\subsection{Derivation}
Now that we have briefly outlined the abstract DPG method, let us apply it to
the convection-diffusion equation. The strong form of the steady
convection-diffusion problem with homogeneous Dirichlet boundary conditions reads
\[
\left\{
\begin{array}[c]{rrl}
\div(\bs\beta u)-\epsilon\del u & =f & \text{in }\Omega\\
u & =0 & \text{on }\Gamma\,,
\end{array}
\right.
\]
where $u$ is the property of interest, $\bs\beta$ is the convection vector,
and $f$ is the source term. Nonhomogeneous Dirichlet and Neumann boundary
conditions are straightforward but would add technicality to the following
discussion. Let us write this as an equivalent system of first
order equations:
\begin{align*}
\div(\bs\beta u-\bs\sigma)&=f\\
\frac{1}{\epsilon}\bs\sigma-\grad u&=\bs0\,.
\end{align*}
If we then multiply the top equation by some scalar test function $v$ and the
bottom equation by some vector-valued test function $\btau$, we can integrate by
parts over each element $K$:
\begin{equation}
\label{eq:preultraweak}
\begin{aligned}
-(\bbeta u-\bfsigma,\nabla v)_K+((\bbeta
u-\bfsigma)\cdot\mathbf{n},v)_{\partial K}&=(f,v)_K\\
\frac{1}{\epsilon}(\bfsigma,\btau)_K+(u,\nabla\cdot\btau)_K
-(u,\tau_n)_{\partial K}&=0\,.
\end{aligned}
\end{equation}
The discontinuous Petrov-Galerkin method refers to the fact that we are using
discontinuous optimal test functions that come from a space differing from the
trial space. It does not specify our choice of trial space. Nevertheless, many
considerations of DPG in the literature (convection-diffusion \cite{DPG6},
linear elasticity \cite{BramwellDemkowiczGopalakrishnanQiu11}, linear
acoustics \cite{DemkowiczGopalakrishnanMugaZitelli12}, Stokes
\cite{DPGStokes}) associate DPG with the so-called ``ultra-weak formulation.''
We will follow the same derivation for the convection-diffusion equation, but
we emphasize that other formulations are available (in particular, the
Primal DPG\cite{PrimalDPG} method presents an alternative with
continuous trial functions). Thus, we seek field variables $u\in L^2(K)$ and
$\bfsigma\in\mb{L^2}(K)$. Mathematically, this leaves their traces on element
boundaries undefined, and in a manner similar to the hybridized discontinuous
Galerkin method, we define new unknowns for trace $\hat u$ and flux $\hat t$.
Applying these definitions to \eqnref{preultraweak} and adding the two
equations together, we arrive at our desired variational problem.

@@

In the following analysis, we neglect the error due to the approximation of optimal test functions.
We follow the classical Brezzi's theory \cite{Brezzi1974,DBB05} for an abstract
mixed problem:
\begin{equation}
\left\{
\begin{array}{lll}
\bfu \in \bfU, p \in Q\\
a(\bfu,\bfw) + c(p,\bfw) & = l(\bfw) & \forall \bfw \in \bfU \\
c(q,\bfu) & = g(q) & \forall q \in Q
\end{array}
\right.
\end{equation}
where $\bfU,Q$ are Hilbert spaces, and $a,c,l,g$ denote the appropriate
bilinear and linear forms. Note that $a(\bfu,\bfw)=b(\bfu,T\bfw)=(T\bfu,T\bfw)_V$ in
the notation from the previous section.

Let function $\bfpsi$ denote the $H(\text{div},\Omega)$ extension of flux $\hat{t}$
that realizes the minimum in the definition of the quotient (minimum energy
extension) norm.
The choice of norm for the Lagrange multipliers $\lambda_K$ is implied
by the quotient norm used for $H^{-1/2}(\Gamma_h)$ and continuity
bound for form $c(p,\bfw)$ representing the constraint:
\begin{equation}
\begin{array}{lll}
| c(\sum_K \lambda_K (1_K,{\bf 0}),(u,\bfsig,\hat{u},\hat{t})) |
& = | \sum_K \lambda_K \langle \hat{t}, 1_K \rangle_{\ptl K} | \\[8pt]
& = | \sum_K \lambda_K \langle v_n , 1_K \rangle_{\ptl K} | \\[8pt]
& = | \sum_K \lambda_K \int_K \text{div} \bfpsi \: 1_K  | \\[8pt]
& \leq \sum_K  \lambda_K || \text{div} \bfpsi ||_{L^2(K)} \mu(K)^{1/2} \\[8pt]
& \leq (\sum_K \mu(K) \lambda_K^2)^{1/2} \: (\sum_K || \text{div} \bfpsi ||_{L^2(K)}^2 )^{1/2} \\[8pt]
& \leq \underbrace{\left(\sum_K \mu(K) \lambda_K^2\right)^{1/2}}_{=: || \bflambda ||} ||
\hat{t} ||_{H^{-1/2}(\Gamma_h)}
\end{array}
\end{equation}
where $\mu(K)$ stands for the area (measure) of element $K$.

@@

Classical $h$-interpolation interpolation error estimate for the lowest error
Raviart-Thomas elements and continuity of operator $R$ imply the stability estimate:
\begin{equation}
\begin{array}{lll}
|| \bfpsi_h || & \leq || \bfpsi_h - \bfpsi || + || \bfpsi ||\\[8pt]
&\leq C h || \bfpsi ||_{H^1} +  || \bfpsi || \\[8pt]
& \leq C || \text{div} \bfpsi || = C (\sum_K \mu(K) \lambda_K^2)^{1/2}
\end{array}
\label{eq:stab}
\end{equation}
Above, $C$ is a generic, mesh independent constant incorporating constant from
the interpolation error estimate and continuity constant of $R$.
Let $\hat{t}$ be the trace of $\bfpsi_h$. We have then,
\begin{equation}
\sup_{\hat{t} \in H^{-1/2}(\Gamma_h)} \frac{|  \sum_K \lambda_K \langle
\hat{t},1_K \rangle_{\ptl K} |}{|| \hat{t} ||_{H^{-1/2}(\Gamma_h)}}
\geq \frac{| \sum_K \lambda_K \int_K \text{div} \bfpsi_h \: 1_K  |}
{|| \bfpsi_h ||_{H(\text{div},\Omega)}}
\geq \frac{1}{C} (\sum_K \mu(K) \lambda_K^2)^{1/2}
\end{equation}
where $C$ is the constant from stability estimate~(\ref{eq:stab}).

Notice that we have considered traces of lowest order Raviart-Thomas elements for the
discretization of flux $\hat{t}$. The inf-sup condition for the lowest order RT spaces implies
automatically the analogous condition for elements of arbitrary order; increasing
space $U$ only makes the discrete inf-sup constant bigger.

\paragraph{Inf Sup in Kernel Condition} is satisfied automatically due to the use of optimal
test functions. First of all, we characterize the ``kernel'' space:
\begin{equation}
\begin{array}{rl}
\bfU_0  := & \{ \bfw \in \bfU \, : \, c(q,\bfw) = 0 \quad \forall q \in Q\} \\[8pt]
 = &\{ (u,\bfsig,\hat{u},\hat{t}) \, : \, \langle \hat{t},1_K \rangle = 0
 \quad \forall K \}
\end{array}
\end{equation}
In other words, the kernel space contains only the equilibrated fluxes.
With $\bfu \in \bfU_0$, we have then:
\begin{equation}
   \sup_{\bfw \in \bfU_0} \frac{| a(\bfu,\bfw) |}{|| \bfw ||_{\bfU} }
   \geq \frac{| b(\bfu,T \bfu) |}{|| \bfu ||}
   = \frac{| b(\bfu,T \bfu) |}{|| T\bfu ||}\frac{||T\bfu||}{||\bfu||}
   = \sup_{(v,\bftau)}\frac{| b((u,\bfsig,\hat{u},\hat{t}), (v,\bftau)) |}{|| (v,\bftau) ||}
   \frac{||T\bfu||}{||\bfu||}
   \geq \gamma^2 || (u,\bfsig,\hat{u},\hat{t}) ||
\end{equation}
where $\gamma$ is the stability constant for the standard continuous DPG formulation.
The first inequality follows as we plug in the definition for $a$ and pick
$\bfw=\bfu$. The second equality is trivial, while the next one follows by definition of the optimal test
functions given through the trial-to-test operator $T$. The finally inequality
springs from the fact that
$\sup_{\bfv}\frac{|b(\bfu,\bfv)|}{||\bfv||}\geq\gamma||\bfu||$ and
$||T\bfu||_V=||R_V^{-1}B\bfu||_V=||B\bfu||_{V'}\geq\gamma||u||$.

@@

Recall the line of analysis leading to the construction of robust
test norms allowing us to bound the $L^2$ error of the field variables by the
energy error, Equation~(\ref{eq:robustBound}). With robust test norms, we have
\begin{equation}
\begin{array}{ll}
   \LRp{|| u - u_h ||^2+ || \bfsig - \bfsig_h ||^2}^{\frac{1}{2}}
& \lesssim || (u - u_h, \bfsig - \bfsig_h, \hat{u} - \hat{u}_h, \hat{t} - \hat{t}_h ||_E \\[8pt]
&= \inf_{(w_h,\bfvarsigma_h,\hat{w},\hat{r}_h)}
|| (u - w_h, \bfsig - \bfvarsigma_h, \hat{u} - \hat{w}_h, \hat{t} - \hat{r}_h ||_E\,.
\end{array}
\end{equation}
The last equality follows from the fact that DPG method delivers the best approximation
error in the energy norm (minimizes the residual). This is no longer true for the
restricted version. So, can we claim robustness in the sense of the inequality above
for the restricted version as well?

@@

\subsection{Robust Test Norms}
The optimal test functions are determined by solving local problems determined
by the choice of test norm. There are several options to consider.
The graph norm \cite{DPGOverview} is one of the most natural norms to consider
as it is derived directly from the adjoint of the problem supplemented with
(possibly scaled) $L^2$ field terms to upgrade it from a semi-norm.
Chan \etal\cite{ChanHeuerThanhDemkowicz2012} derived a more robust alternative
norm for convection diffusion (dubbed the robust norm). We recently
developed a modification of the robust norm that appears to produce better
results in the presence of singularities. In this section we illustrate the
singularity issues with the Laplace equation, offer possible explanations for
the phenomena observed, and propose a modification of the robust test norm,
which we demonstrate eliminates the issues observed in numerical experiments.

\subsubsection{A Model Problem}
\seclab{sec:confusionPlate}
We begin by first examining a different problem than convection-diffusion.
We look at admissible solutions for the homogeneous Laplace's equation
(convection diffusion with $\bbeta=\bs 0$) over
the $y > 0$ half-plane under boundary conditions
\begin{align*}
u &= 0 \text{ on } x > 0\\
\pd{u}{n} &= 0 \text{ on } x < 0.
\end{align*}
Let us consider the 2D case - a simple separation of variables argument in
polar coordinates shows that the solution is of the form
\[
u(r,\theta) = \sum_{n=0}^\infty R_n(r) \sin(\lambda_n \theta),
\]
where $\lambda_n = n + \frac{1}{2}$, and $R_n(r) = C_{1,n}x^{\lambda_n} +
C_{2,n}x^{-\lambda_n}$.  By requiring $u(0,\theta) < \infty$, we have $R_n(r)
= C_n x^{\lambda_n}$.  We have now that solutions to this problem include $u$
of the form
\[
u = \sum_{n=0} x^{n+\frac{1}{2}} \sin\LRp{\LRp{n+\frac{1}{2}}\theta}.
\]
Note that even in the lowest-order term, the gradient of $u$ displays a
singularity at $r = 0$. It is well known that, for smooth boundary data,
solutions to Laplace’s equation can be decomposed into the linear combination
of smooth and singular contributions; the above analysis implies that, when
boundary conditions change from Dirichlet to Neumann on the half-plane, the
Laplace’s equation will always develop a singularity in the stresses.

@@

\subsubsection{A Modification of the Robust Test Norm}

@@

\section{Numerical Experiments}
In \ref{sec:problemDescriptions} we define each numerical experiment, and in
\ref{sec:problemAnalysis} we discuss the solution properties in general. We
solve with second order field variables and flux ($u$, $\bfsigma$, and $\hat
t$), third order traces ($\hat t$), and fifth order test functions ($v$ and
$\btau$).

>>>>

\documentclass[letterpaper]{article}
%TODO: check for periods after equations
%TODO: bold face proper variables
%TODO: Make sure titles have consistent case
\usepackage[margin=1.0in]{geometry}
\usepackage{authblk}
\usepackage{amsmath, amssymb, amsthm, mathtools}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{color}
\usepackage{subcaption}
\usepackage{morefloats}
\usepackage{pgf,tikz}
\usetikzlibrary{arrows}

@@

\begin{abstract}
We develop a locally conservative formulation of the discontinuous
Petrov-Galerkin finite element method (DPG) for convection-diffusion type
problems using Lagrange multipliers to exactly enforce conservation over each
element. We provide a proof of convergence as well as extensive numerical
experiments showing that the method is indeed locally conservative. We also
show that standard DPG, while not guaranteed to be conservative, is nearly
conservative for many of the benchmarks considered. The new method preserves many of
the attractive features of DPG, but turns the normally symmetric
positive-difinite DPG system into a saddlepoint problem.
\end{abstract}

\section{Introduction}
The discontinuous Petrov-Galerkin (DPG) method with optimal test functions has
been under active development for convection-diffusion type systems
\cite{DPG1, DPG2, DPG3, DPG5, DemkowiczHeuer, ChanHeuerThanhDemkowicz2012,
MoroNguyenPeraire11}. In this paper, we develop a theory for the locally
conservative formulation of DPG for convection-diffusion type equations
(including Burgers' Equation and Stokes flow) and supplement this with
extensive numerical results.

@@

The discontinuous Petrov-Galerkin finite element method has been described as
least squares finite elements with a twist. The key difference is that least
squares methods seek to minimize the residual of the solution in the $L^2$
norm, while DPG seeks the minimization in a dual norm realized through the
inverse Riesz map. Exact mass conservation has been an issue that has long plagued
least squares finite elements. Several approaches have been
used to try to adress this. Bochev \etal\cite{Bochev2010} accomplish local
conservation by using a pointwise divergence free velocity space in the Stokes
formulation.  Chang and Nelson\cite{ChangNelson1997} developed the
\emph{restricted LSFEM}\cite{ChangNelson1997} by augmenting the least squares
equations with Lagrange multipliers explicitly enforcing mass conservation
element-wise. Our conservative formulation of DPG takes a similar approach and
both methods similarly share the disadvantage of transforming a minimization method to a
saddle-point problem. In the interest of crediting Chang and Nelson's
restricted LSFEM, we call the following locally conservative DPG method the
restricted DPG method (RDPG).

\subsection{DPG is a Minimum Residual Method}
Roberts \etal presents a brief history and derivation of DPG with optimal test functions in
\cite{DPGStokes}. We follow his derivation of the standard DPG method as a
minumum residual method. Let $U$ be the trial Hilbert space and $V$ the test
Hilbert space for a well-posed variational problem $b(u,v)=l(v)$. In operator
form this is $Bu=l$, where $B:U\rightarrow V'$ and $\LRa{Bu,v}=b(u,v)$. We seek to minimize the
residual for the discrete space $U_h\subset U$:
\begin{equation}
u_h=\argmin_{u_h\in U_h}\frac{1}{2}\norm{Bu_h-l}^2_{V'}\,.
\label{minresidual}
\end{equation}
Recalling that the Riesz operator $R_V:V\rightarrow V'$ is an isometry defined
by
\[
\LRa{R_Vv,\delta v}=\LRp{v,\delta v}_V,\quad\forall\delta v\in V,
\]
we can use the Riesz inverse to minimize in the $V$-norm rather than its dual:
\begin{equation}
\frac{1}{2}\norm{Bu_h-l}^2_{V'}=\frac{1}{2}\norm{R_V^{-1}(Bu_h-l)}^2_V
=\frac{1}{2}\LRp{R_V^{-1}(Bu_h-l),R_V^{-1}(Bu_h-l)}_V\,.
\label{eq:rieszapplied}
\end{equation}
The first order optimality condition for \eqnref{rieszapplied} requires
the G\^ateaux derivative to be zero in all directions $\delta u \in
U_h$, i.e.,
\[
\left(R_V^{-1}(Bu_h-l),R_V^{-1}B\delta u\right)_V = 0, \quad \forall \delta u \in U.
\]
By definition of the Riesz operator, this is equivalent to
\begin{equation}
\LRa{Bu_h-l,R_V^{-1}B\delta u_h}=0\quad\forall\delta u_h\in U_h\,.
\label{eq:DPGbilinearform}
\end{equation}
Now, we can identify $v_{\delta u_h}\coloneqq R_V^{-1}B\delta u_h$ as the
optimal test function for trial function $\delta u_h$. Define $T:=R_V^{-1}B:U_h\rightarrow V$ as the trial-to-test operator. Now we can rewrite
\eqnref{DPGbilinearform} as
\begin{equation}
b(u_h,v_{\delta u_h})=l(v_{\delta u_h}).
\label{eq:DPGmethod}
\end{equation}
The DPG method then is to solve \eqnref{DPGmethod} with optimal test functions
$v_{\delta u_h}\in V$ that solve the auxiliary problem
\begin{equation}
\LRp{v_{\delta u_h},\delta v}_V=\LRa{R_Vv_{\delta u_h},\delta v}
=\LRa{B\delta u_h,\delta v}=b(\delta u_h,\delta v)\,,\quad\forall\delta v\in V.
\label{eq:optimaltestproblem}
\end{equation}
Using a continuous test basis would result in a global solve for every optimal
test function. Therefore DPG uses a discontinuous test basis which makes each
solve element-local and much more computationally tractable. Of course,
\eqnref{optimaltestproblem} still requires the inversion of the
infinite-dimensional Riesz map, but approximating $V$ by a finite
dimensional space, $V_h$, which is of a higher polynomial degree than $U_h$ (hence
``enriched space'') works well in practice.

@@

\subsection{Derivation}
Now that we have briefly outlined the abstract DPG method, let us apply it to
the convection-diffusion equation. The strong form of the steady
convection-diffusion problem with homogeneous Dirichlet boundary conditions reads
\[
\left\{
\begin{array}[c]{rrl}
\div(\bs\beta u)-\epsilon\del u & =f & \text{in }\Omega\\
u & =0 & \text{on }\Gamma\,,
\end{array}
\right.
\]
where $u$ is the property of interest, $\bs\beta$ is the convection vector,
and $f$ is the source term. Nonhomogeneous Dirichlet and Neumann boundary
conditions are straightforward but would add technicality to the following
discussion. Let us write this as an equivalent system of first
order equations:
\begin{align*}
\div(\bs\beta u-\bs\sigma)&=f\\
\frac{1}{\epsilon}\bs\sigma-\grad u&=\bs0\,.
\end{align*}
If we then multiply the first equation by some scalar test function $v$ and the
bottom equation by some vector-valued test function $\btau$, we can integrate by
parts over each element $K$:
\begin{equation}
\label{eq:preultraweak}
\begin{aligned}
-(\bbeta u-\bfsigma,\nabla v)_K+((\bbeta
u-\bfsigma)\cdot\mathbf{n},v)_{\partial K}&=(f,v)_K\\
\frac{1}{\epsilon}(\bfsigma,\btau)_K+(u,\nabla\cdot\btau)_K
-(u,\tau_n)_{\partial K}&=0\,.
\end{aligned}
\end{equation}
The discontinuous Petrov-Galerkin method refers to the fact that we are using
discontinuous optimal test functions that come from a space differing from the
trial space. It does not specify our choice of trial space. Nevertheless, many
versions of DPG in the literature (convection-diffusion \cite{DPG6},
linear elasticity \cite{BramwellDemkowiczGopalakrishnanQiu11}, linear
acoustics \cite{DemkowiczGopalakrishnanMugaZitelli12}, Stokes
\cite{DPGStokes}) associate DPG with the so-called ``ultra-weak formulation.''
We will follow the same derivation for the convection-diffusion equation, but
we emphasize that other formulations are available (in particular, the
Primal DPG\cite{PrimalDPG} method presents an alternative with
continuous trial functions). Thus, we seek field variables $u\in L^2(K)$ and
$\bfsigma\in\mb{L^2}(K)$. Mathematically, this leaves their traces on element
boundaries undefined, and in a manner similar to the hybridized discontinuous
Galerkin method, we define new unknowns for trace $\hat u$ and flux $\hat t$.
Applying these definitions to \eqnref{preultraweak} and adding the two
equations together, we arrive at our desired variational problem.

@@

In the following analysis, we neglect the error due to the approximation of optimal test functions.
We follow the classical Brezzi's theory \cite{Brezzi1974,DBB05} for an abstract
mixed problem:
\begin{equation}
\left\{
\begin{array}{lll}
\bfu \in \bfU, p \in Q\\
a(\bfu,\bfw) + c(p,\bfw) & = l(\bfw) & \forall \bfw \in \bfU \\
c(q,\bfu) & = g(q) & \forall q \in Q
\end{array}
\right.
\end{equation}
where $\bfU,Q$ are Hilbert spaces, and $a,c,l,q$ denote the appropriate
bilinear and linear forms. Note that $a(\bfu,\bfw)=b(\bfu,T\bfw)=(T\bfu,T\bfw)_V$ in
the notation from the previous section.

Let function $\bfpsi$ denote the $H(\text{div},\Omega)$ extension of flux $\hat{t}$
that realizes the minimum in the definition of the quotient (minimum energy
extension) norm.
The choice of norm for the Lagrange multipliers $\lambda_K$ is implied
by the quotient norm used for $H^{-1/2}(\Gamma_h)$ and continuity
bound for form $c(p,\bfw)$ representing the constraint:
\begin{equation}
\begin{array}{lll}
| c(\sum_K \lambda_K (1_K,{\bf 0}),(u,\bfsig,\hat{u},\hat{t})) |
& = | \sum_K \lambda_K \langle \hat{t}, 1_K \rangle_{\ptl K} | \\[8pt]
& = | \sum_K \lambda_K \langle v_n , 1_K \rangle_{\ptl K} | \\[8pt]
& = | \sum_K \lambda_K \int_K \text{div} \bfpsi \: 1_K  | \\[8pt]
& \leq \sum_K  \lambda_K || \text{div} \bfpsi ||_{L^2(K)} \mu(K)^{1/2} \\[8pt]
& \leq (\sum_K \mu(K) \lambda_K^2)^{1/2} \: (\sum_K || \text{div} \bfpsi ||_{L^2(K)}^2 )^{1/2} \\[8pt]
& \leq \underbrace{(\sum_K \mu(K) \lambda_K^2)^{1/2}}_{=: || \bflambda ||} ||
\hat{t} ||_{H^{-1/2}(\Gamma_h)}
\end{array}
\end{equation}
where $\mu(K)$ stands for the area (measure) of element $K$.

@@

Classical $h$-interpolation interpolation error estimates for the lowest error
Raviart-Thomas elements and continuity of operator $R$ imply the stability estimate:
\begin{equation}
\begin{array}{lll}
|| \bfpsi_h || & \leq || \bfpsi_h - \bfpsi || + || \bfpsi ||\\[8pt]
&\leq C h || \bfpsi ||_{H^1} +  || \bfpsi || \\[8pt]
& \leq C || \text{div} \bfpsi || = C (\sum_K \mu(K) \lambda_K^2)^{1/2}
\end{array}
\label{eq:stab}
\end{equation}
Above, $C$ is a generic, mesh independent constant incorporating constant from
the interpolation error estimate and continuity constant of $R$.
Let $\hat{t}$ be now the trace of $\bfpsi_h$. We have then,
\begin{equation}
\sup_{\hat{t} \in H^{-1/2}(\Gamma_h)} \frac{|  \sum_K \lambda_K \langle
\hat{t},1_K \rangle_{\ptl K} |}{|| \hat{t} ||_{H^{-1/2}(\Gamma_h)}}
\geq \frac{| \sum_K \lambda_K \int_K \text{div} \bfpsi_h \: 1_K  |}
{|| \bfpsi_h ||_{H(\text{div},\Omega)}}
\geq \frac{1}{C} (\sum_K \mu(K) \lambda_K^2)^{1/2}
\end{equation}
where $C$ is the constant from stability estimate~(\ref{eq:stab}).

Notice that we have considered traces of lowest order Raviart-Thomas elements
for the discretization of flux $\hat{t}$. The inf-sup condition for the lowest
order RT spaces implies automatically the analogous condition for elements of
arbitrary order; increasing the dimension of space $U$ only makes the discrete
inf-sup constant bigger.

\paragraph{Inf Sup in Kernel Condition} is satisfied automatically due to the use of optimal
test functions. First of all, we characterize the ``kernel'' space:
\begin{equation}
\begin{array}{rl}
\bfU_0  := & \{ \bfw \in \bfU \, : \, c(q,\bfw) = 0 \quad \forall q \in Q\} \\[8pt]
 = &\{ (u,\bfsig,\hat{u},\hat{t}) \, : \, \langle \hat{t},1_K \rangle = 0
 \quad \forall K \}
\end{array}
\end{equation}
In other words, the kernel space contains only the equilibriated fluxes.
With $\bfu \in \bfU_0$, we have then:
\begin{equation}
   \sup_{\bfw \in \bfU_0} \frac{| a(\bfu,\bfw) |}{|| \bfw ||_{\bfU} }
   \geq \frac{| b(\bfu,T \bfu) |}{|| \bfu ||}
   = \frac{| b(\bfu,T \bfu) |}{|| T\bfu ||}\frac{||T\bfu||}{||\bfu||}
   = \sup_{(v,\bftau)}\frac{| b((u,\bfsig,\hat{u},\hat{t}), (v,\bftau)) |}{|| (v,\bftau) ||}
   \frac{||T\bfu||}{||\bfu||}
   \geq \gamma^2 || (u,\bfsig,\hat{u},\hat{t}) ||
\end{equation}
where $\gamma$ is the stability constant for the standard continuous DPG formulation.
The first inequality follows as we plug in the definition for $a$ and pick
$\bfw=\bfu$. The second equality is trivial, while the next one follows by definition of the optimal test
functions given through the trial-to-test operator $T$. The finally inequality
springs from the fact that
$\sup_{\bfv}\frac{|b(\bfu,\bfv)|}{||\bfv||}\geq\gamma||\bfu||$ and
$||T\bfu||_V=||R_V^{-1}B\bfu||_V=||B\bfu||_{V'}\geq\gamma||u||$.

@@

Recall the line of analysis leading to the construction of robust
test norms allowing us to bound the $L^2$ error of the field variables by the
energy error, Equation \ref{eq:robustBound}. With robust test norms, we have
\begin{equation}
\begin{array}{ll}
   \LRp{|| u - u_h ||^2+ || \bfsig - \bfsig_h ||^2}^{\frac{1}{2}}
& \lesssim || (u - u_h, \bfsig - \bfsig_h, \hat{u} - \hat{u}_h, \hat{t} - \hat{t}_h ||_E \\[8pt]
&= \inf_{(w_h,\bfvarsigma_h,\hat{w},\hat{r}_h)}
|| (u - w_h, \bfsig - \bfvarsigma_h, \hat{u} - \hat{w}_h, \hat{t} - \hat{r}_h ||_E\,.
\end{array}
\end{equation}
The last equality follows from the fact that DPG method delivers the best approximation
error in the energy norm (minimizes the residual). This is no longer true for the
restricted version. So, can we claim robustness in the sense of the inequality above
for the restricted version as well?

@@

\subsection{Robust test norms}
The optimal test functions are determined by solving local problems determined
by the choice of test norm. There are several options to consider.
The graph norm \cite{DPGOverview} is one of the most natural norms to consider
as it is derived directly from the adjoint of the problem supplemented with
(possibly scaled) $L^2$ field terms to upgrade it from a semi-norm.
Chan \etal\cite{ChanHeuerThanhDemkowicz2012} derived a more robust alternative
norm for convection diffusion (dubbed the robust norm). We recently
developed a modification of the robust norm that appears to produce better
results in the presence of singularities. In this section we illustrate the
singularity issues with the Laplace equation, offer possible explanations for
the phenomena observed, and propose a modification of the robust test norm,
which we demonstrate eliminates the issues observed in numerical experiments.

\subsubsection{A model problem}
\seclab{sec:confusionPlate}
We begin by first examining a different problem than convection-diffusion.
We look at admissible solutions for the homogeneous Laplace's equation
(convection diffusion with $\bbeta=\bs 0$) over
the $y > 0$ half-plane under boundary conditions
\begin{align*}
u &= 0 \text{ on } x > 0\\
\pd{u}{n} &= 0 \text{ on } x < 0.
\end{align*}
Let us consider the 2D case - a simple separation of variables argument in
polar coordinates shows that the solution is of the form
\[
u(r,\theta) = \sum_{n=0}^\infty R_n(r) \sin(\lambda_n \theta),
\]
where $\lambda_n = n + \frac{1}{2}$, and $R_n(r) = C_{1,n}x^{\lambda_n} +
C_{2,n}x^{-\lambda_n}$.  By requiring $u(0,\theta) < \infty$, we have $R_n(r)
= C_n x^{\lambda_n}$.  We have now that solutions to this problem include $u$
of the form
\[
u = \sum_{n=0} x^{n+\frac{1}{2}} \sin\LRp{\LRp{n+\frac{1}{2}}\theta}.
\]
Note that in the lowest-order term, the gradient of $u$ displays a
singularity at $r = 0$. It is well known that, for smooth boundary data,
solutions to Laplace’s equation can be decomposed into the linear combination
of smooth and singular contributions; the above analysis implies that, when
boundary conditions change from Dirichlet to Neumann on the half-plane, the
Laplace’s equation will always develop a singularity in the stresses.

@@

\subsubsection{A modification of the robust test norm}

@@

\section{Numerical Experiments}
In \ref{sec:problemDescriptions} we define each numerical experiment, and in
\ref{sec:problemAnalysis} we discuss the solution properties in general. We
solve with second order field variables and flux ($u$, $\bfsigma$, and $\hat
t$), third order traces ($\hat u$), and fifth order test functions ($v$ and
$\btau$).
